The LSST DESC 3x2pt Tomography Optimization Challenge

### 2021-03-12 paper/paper.tex
add hline in app tables

These methods are described in full in \autoref{app:methods}.  Briefly, they were (with the submitting authors listed in brackets):
% \setlength{\itemsep}{0pt}%
\begin{description*}
	\item[\textsc{Random}] the random assignment control method (organizers)
	\item[\textsc{IBandOnly}] the i-band only control method (organizers)
	\item[\textsc{RandomForest}] a collection of individual decision trees (organizers)
	\item[\textsc{LSTM}] a Long Short-Term Memory neural network (CRB, BMOF, GT, ESC, EJG)
	\item[\textsc{AutokerasLSTM}] an automaticaly configured LSTM variant\footnote{A technical problem, involving conflicting versions of libraries and incompatible hardware, meant that AutoKerasLSTM could not be run in the final challenge.} (CRB, BMOF, GT, ESC, EJG)
	\item[\textsc{LGBM}] a Light Gradient Boosted Machine, a histogram-based tree algorthm (CRB, BMOF, GT, ESC, EJG)
	\item[\textsc{TCN}] a Temporal Convolution Network (CRB, BMOF, GT, ESC, EJG)
	\item[\textsc{CNN}] a Convolutional Neural Network (CRB, BMOF, GT, ESC, EJG)
	\item[\textsc{Ensemble}] a combination of three other network methods (CRB, BMOF, GT, ESC, EJG)
	\item[\textsc{SimpleSOM}] a Self-Organizing Map (AHW)
	\item[\textsc{PQNLD}] an extension of {\sc SimpleSom} combining with template-based redshift estimation (AHW)
	\item[\textsc{ComplexSOM}] an extension of {\sc SimpleSom} with an additional assignment optimization (DA, AHW, AN, EG, AB)
	\item[\textsc{GPzBinning}] a Gaussian Process redshift estimator and binner (PH)
	\item[\textsc{JaxCNN/JaxResnet}] two related CNN-based bin edge optimizers (AP)
	\item[\textsc{NeuralNetwork1/2}] a dense network optimizing bin assignment for two metrics (AP)
	\item[\textsc{PCACluster}] a Principal Component Analysis of fluxes then optimized clustering (DG)
	\item[\textsc{Utopia}] a nearest-neighbour method, optimised in the limit of large representative training sets (AHW)
	\item[\textsc{ZotBin/ZotNet}] two related neural network methods with a preprocesing normalizing flow (DK)
	\item[\textsc{FFNN}] a feed-forward neural network (EN)
	\item[\textsc{FunBins}] a random forest with various nominal bin edge selection (AB)
	\item[\textsc{MLPQNA}] a multi-layer perceptron (SC and MB)
	\item[\textsc{Stacked Generalization}] an combination of gradient tree boosting and logistic regression (JEC)
\end{description*}

@@

Among methods with fixed bins, those using equal numbers in each tend to plateau at scores of
120-140 in Table \ref{tab:cosmodc2}, suggesting that a range of machine learning models can
effectively classify by bin, at least in this case of representative, extensive training
information.  Notably, this is the case even for the UTOPIA entry, which uses the nearest 
training set neighbour to assign bins and thus is explicitly optimised
for this idealised scenario; it should give an upper limit when using the same fixed bins.
That other methods with the same nominal bin choice acheive scores close to it once again
confirms that many entries are close to the best possible score for this approach.

@@

In this way UTOPIA is the simplest possible algorithm (utilising all available
photometric bands) that one can  implement for tomography. It is also a method
whose performance is optimal in the limit of an infinitely large,  perfectly
representative training sample; i.e. the limit in which this challenge operates.
As the requirements of perfect  training-sample representivity and large size
are violated, UTOPIA ought to assign redshifts to reference  galaxies from
increasingly distant regions of the multi-dimensional magnitude-space, resulting
in poorer (i.e. unreliable) tomographic binning.  Therefore, UTOPIA acts as a warning against
over-interpretation of the tomographic challenge results, as success in the
challenge  need not translate to optimal-performance/usefulness under realistic
survey conditions. Instead, the challenge results should be interpreted
holistically, by comparing the performance between different classes of methods,
rather than individuals. The authors have endeavoured to do so in the conclusions presented here.

@@

\section{Full 9-bin results table}
Tables \ref{tab:full_dc2} and \ref{tab:full_buzz} show all calculated metrics when generating nine tomographic 
bins for CosmoDC2 and Buzzard respectively. The \textit{ww} columns
show weak-lensing metrics, \textit{gg} show galaxy clustering, and \textit{3x} the full 3x2pt metric.

As in table \ref{tab:cosmodc2}, some values are missing due to failure or time-out of the method (*) or pathological bin assignments (-).

>>>>

These methods are described in full in \autoref{app:methods}.  Briefly, they were (with the submitting authors listed in brackets):
% \setlength{\itemsep}{0pt}%
\begin{description*}
	\item[\textsc{Random}] the random assignment control method (organizers)
	\item[\textsc{IBandOnly}] the i-band only control method (organizers)
	\item[\textsc{RandomForest}] a collection of individual decision trees (organizers)
	\item[\textsc{LSTM}] a Long Short-Term Memory neural network (CRB, BMOF, GT, ESC, EJG)
	\item[\textsc{AutokerasLSTM}] an automaticaly configured LSTM variant\footnote{A technical problem, involving conflicting versions of libraries and incompatible hardware, meant that AutoKerasLSTM could not be run in the final challenge.} (CRB, BMOF, GT, ESC, EJG)
	\item[\textsc{LGBM}] a Light Gradient Boosted Machine, a histogram-based tree algorthm (CRB, BMOF, GT, ESC, EJG)
	\item[\textsc{TCN}] a Temporal Convolution Network (CRB, BMOF, GT, ESC, EJG)
	\item[\textsc{CNN}] a Convolutional Neural Network (CRB, BMOF, GT, ESC, EJG)
	\item[\textsc{Ensemble}] a combination of three other network methods (CRB, BMOF, GT, ESC, EJG)
	\item[\textsc{SimpleSOM}] a Self-Organizing Map (AHW)
	\item[\textsc{PQNLD}] an extension of {\sc SimpleSom} combining with template-based redshift estimation (AHW)
	\item[\textsc{ComplexSOM}] an extension of {\sc SimpleSom} with an additional assignment optimization (DA, AHW, AN, EG, AB)
	\item[\textsc{GPzBinning}] a Gaussian Process redshift estimator and binner (PH)
	\item[\textsc{JaxCNN/JaxResnet}] two related CNN-based bin edge optimizers (AP)
	\item[\textsc{NeuralNetwork1/2}] a dense network optimizing bin assignment for two metrics (AP)
	\item[\textsc{PCACluster}] a Principal Component Analysis of fluxes then optimized clustering (DG)
	\item[\textsc{UTOPIA}] a nearest-neighbour method, optimised in the limit of large representative training sets (AHW)
	\item[\textsc{ZotBin/ZotNet}] two related neural network methods with a preprocesing normalizing flow (DK)
	\item[\textsc{FFNN}] a feed-forward neural network (EN)
	\item[\textsc{FunBins}] a random forest with various nominal bin edge selection (AB)
	\item[\textsc{MLPQNA}] a multi-layer perceptron (SC and MB)
	\item[\textsc{Stacked Generalization}] an combination of gradient tree boosting and logistic regression (JEC)
\end{description*}

@@

Among methods with fixed bins, those using equal numbers in each tend to plateau at scores of
120-140 in Table \ref{tab:cosmodc2}, suggesting that a range of machine learning models can
effectively classify by bin, at least in this case of representative, extensive training
information.  Notably, this is the case even for the \textsc{UTOPIA} entry, which uses the nearest 
training set neighbour to assign bins and thus is explicitly optimised
for this idealised scenario; it should give an upper limit when using the same fixed bins.
That other methods with the same nominal bin choice acheive scores close to it once again
confirms that many entries are close to the best possible score for this approach.

@@

In this way \textsc{UTOPIA} is the simplest possible algorithm (utilising all available
photometric bands) that one can  implement for tomography. It is also a method
whose performance is optimal in the limit of an infinitely large,  perfectly
representative training sample; i.e. the limit in which this challenge operates.
As the requirements of perfect  training-sample representivity and large size
are violated, \textsc{UTOPIA} ought to assign redshifts to reference  galaxies from
increasingly distant regions of the multi-dimensional magnitude-space, resulting
in poorer (i.e. unreliable) tomographic binning.  Therefore, \textsc{UTOPIA} acts as a warning against
over-interpretation of the tomographic challenge results, as success in the
challenge  need not translate to optimal-performance/usefulness under realistic
survey conditions. Instead, the challenge results should be interpreted
holistically, by comparing the performance between different classes of methods,
rather than individuals. The authors have endeavoured to do so in the conclusions presented here.

@@

\section{Full 9-bin results table}
Tables \ref{tab:full_dc2} and \ref{tab:full_buzz} show all calculated metrics when generating nine tomographic 
bins for CosmoDC2 and Buzzard respectively. The \textit{ww} columns
show weak-lensing metrics, \textit{gg} show galaxy clustering, and \textit{3x} the full 3x2pt metric.
Methods above the horizontal line trained bin edges; methods below used fixed fiducial ones, though in some
cases did some hand-tuning of them before submission.

As in table \ref{tab:cosmodc2}, some values are missing due to failure or time-out of the method (*) or 
pathological bin assignments (-).  Failure rates were higher for the 9-bin runs than for the lower bin counts.
